---
output: 
  github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->  

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(collapse = TRUE)

library(tidyverse)

```

# S2MET_StabilityMapping

## Description

This repository contains information and code for replicating the analyses described in the paper:

Article Title: *Genomewide association and prediction of phenotypic stability in barley*  
Journal: *The Plant Genome* (under review)  
Authors: Jeffrey L. Neyhart, Lucia Gutierrez, and Kevin P. Smith  
Article doi: TBD



## Navigation

### Data

Data used in this study are available from the [Triticeae Toolbox](https://triticeaetoolbox.org/barley). See [this README](`r file.path(gsub(pattern = "\\.git", replacement = "", x = git2r::remote_url()), "tree/master/Data")`) for instructions on accessing this data.

### Code

Scripts used to complete the analysis and generate figures outlined in the article above are available in the "Scripts" subfolder. See [this README](`r file.path(gsub(pattern = "\\.git", replacement = "", x = git2r::remote_url()), "tree/master/Scripts")`) for information on the scripts and their intended execution order.

Three scripts in this directory are used by all other scripts:

1. `startub.R` - loads packages, creates directory links, and loads data.
2. `functions.R` - loads additional functions into the environment. 

The `06_figures.R` script in the "Scripts" subfolder produces the figures found in the paper.


## Software/package versions

*R* version `r paste0(version$major, ".", version$minor)` was used to complete analyses in this project.

The following packages (with version information) were used to complete this project:


```{r version, warning=FALSE, message=FALSE, echo=FALSE}

# Scan the source function and all scripts for packages
packages <- list.files(path = getwd(), full.names = TRUE, pattern = ".R$", recursive = TRUE) %>% 
  str_subset(string = ., pattern = "archive", negate = TRUE) %>%
  map(readLines) %>% 
  map(~str_subset(., "library\\(|require\\(")) %>% 
  unlist() %>% 
  str_subset(., "#", negate = TRUE) %>%
  str_trim() %>%
  str_extract("\\(.*\\)") %>% 
  str_remove_all("\\(|\\)") %>% 
  unique()

# get descriptions of each package; create a tibble
package_info <- installed.packages() %>% 
  as_tibble() %>% 
  filter(Package %in% packages) %>% 
  select(package = Package, version = Version) %>%
  arrange(package)

# Output a table
knitr::kable(package_info)


```