---
title: "S2MET Mapping Phenotypic Data Summary"
output: html_notebook
bibliography: C:/Users/Jeff/Documents/Literature/library.bib
---

## Introduction

This notebook will provide some phenotyping data summaries for the S2MET Mapping project. It will include:

1. Basic model for g + e + gxe
2. Heritability estimates
3. Correlations among environments

```{r prep}

library(tidyverse)
library(broom)
library(stringr)
library(readxl)
library(modelr)
library(pbr)
library(rrBLUP)
library(ggridges)

# Project and other directories
source("C:/Users/Jeff/Google Drive/Barley Lab/Projects/S2MET_Mapping/source.R")

tp_geno <- tp_geno_multi

load(file.path(geno_dir, "S2_genos_hmp.RData"))

# Filter environments for those in which the TP was observed
S2_MET_BLUEs_use <- S2_MET_BLUEs %>% 
  filter(line_name %in% tp_geno,
         trait != "TestWeight")

```


## Results

### Basic Summaries



Look at the number of lines per environment

```{r line.summ}

# Find the total number of possible line x environment combinations and find
# the proportion that are observed for each trait
(prob_observed <- S2_MET_BLUEs_use %>% 
  distinct(trait, environment, line_name) %>%
  mutate(observed = TRUE) %>% 
  group_by(trait) %>%
  complete(environment, line_name, fill = list(observed = FALSE)) %>%
  # group_by(trait) %>%
  summarize(prop_obs = mean(observed)))


```

Interpretation: of the possible line x environment combinations in which at least one line was observed, the above is the proportion of the lines that were observed.






Visualization of distributions

```{r pheno.viz}

# Sort on grain yield environmental mean
env_order <- S2_MET_BLUEs_use %>% 
  mutate_if(is.character, as.factor) %>% 
  group_by(environment, trait) %>% 
  mutate(env_mean = mean(value, na.rm = TRUE)) %>% 
  filter(trait == "GrainYield") %>% 
  complete(environment) %>%
  arrange(env_mean) %>%
  pull(environment) %>% 
  unique()

(g_met_dist <- S2_MET_BLUEs_use %>%
  filter(trait != "TestWeight") %>%
  mutate(environment = parse_factor(environment, levels = env_order)) %>%
  ggplot(aes(x = value, y = environment, fill = environment)) +
  geom_density_ridges() +
  facet_grid(. ~ trait, scales = "free_x") +
  scale_fill_discrete(guide = FALSE) +
  ylab("Environment") +
  xlab("") +
  labs(title = "Trait Distributions in All Environments") +
  theme_bw() )

# Sort

# Save it
save_file <- file.path(fig_dir, "met_trait_dist.jpg")
ggsave(filename = save_file, plot = g_met_dist, width = 7, height = 5)


```



### Heritability



The heritability will be calcuted for each trait across all environments by fitting the following model:

$$
y_{ij} = \mu + g_i + t_j + (gt)_{ij} + \epsilon_{ij}
$$

where $y_{ij}$ is the adjusted mean of the *i*th genotype in the *j*th environment, $\mu$ is the grand mean, $g_i$ is the random effect of the *i*th genotype, $t_j$ is the fixed effect of the $j$th environment, $(gt)_{ij}$ is the random effect of the interaction between the $i$th genotype and the $j$th environment, and $\epsilon_{ij}$ is the error associated with the observation.

We will assume the random effects are distributed such that $g \sim N(0, \sigma^2_g)$, $(gt) \sim N(0, \sigma^2_{(gt)})$, and $\epsilon \sim N(0, \mathbf{R}\sigma^2_\epsilon)$, where $\mathbf{R}$ is a diagonal matrix with elements equal to the inverse of the variances of the adjusted genotype means.



```{r herit.across}

# Use a different optimizer
library(optimx)

# Group by trait and fit the multi-environment model
stage_two_model_fits <- S2_MET_BLUEs_use %>% 
  group_by(trait) %>%
  # mutate(value = scale(value)) %>% # Scale and center the observations
  rename(env = environment) %>% # Rename environment
  do({
    # Extract the df
    df <- .
    
    # List of models
    models <- formulas(~ value,
                      base = ~ 1 + (1|env) + (1|line_name),
                      full = add_predictors(base, ~ (1 | line_name:env)))
    
    # Lmer control
    lmer_control <- lmerControl(check.nobs.vs.nlev = "ignore", check.nobs.vs.nRE = "ignore",
                                optimizer ='optimx', optCtrl=list(method='nlminb'))
    # Get the weights
    wts <- df$std_error^2
    
    # Fit the models
    fits <- models %>% 
      map(~lmer(formula = ., data = df, control = lmer_control, weights = wts))
    
    # Table of lines by environments (i.e. plots)
    plot_table <- xtabs(formula = ~ line_name + env, data = df)
    
    # Find the number of environments
    n_e <- plot_table %>%
      ifelse(. > 1, 1, .) %>%
      rowSums() %>% 
      harm_mean()
    
    # Now replicates
    n_r <- plot_table %>% 
      harm_mean()
    
    # Return data_frame
    data_frame(fits = list(fits), n_e = n_e, n_r = n_r) })
    

```


First test for significance of the $gt$ term using a liklihood ratio test

The likelihood ratio is calculated as

$$
LR = -2[\log(L_{Red}) - \log(L_{Full})]
$$

The $LR$ is $\chi^2$ distributed with 1 degree of freedom (since we dropped one variance component to get the reduced model). The hypothesis test is $H_0: \sigma^2_i = 0$ and $H_A: \sigma^2_i > 0$

```{r lrt}

stage_two_lrt <- stage_two_model_fits %>% 
  mutate(llik = map(fits, ~map_df(., logLik))) %>% 
  unnest(llik) %>%
  select(-fits:-n_r) %>%
  mutate(lr = -2 * (base - full),
         p_value = pchisq(q = lr, df = 1, lower.tail = FALSE) / 2)


stage_two_lrt

```


The interaction term is significant for all traits



The broad-sense heritability is calculated on a entry-mean basis according to the formula

$$
H = \frac{ \sigma^2_G }{ \sigma^2_G + \frac{\sigma^2_{GE}}{r} + \frac{\sigma^2_R}{er} }
$$

for overall heritability calculations, and 

$$
H = \frac{ \sigma^2_G }{ \sigma^2_G  + \frac{\sigma^2_R}{r} }
$$


where $e$ is the number of environments and $r$ is the number of replicates. In the case of an augmented design, where the number of replicates is uneven, the following formula is used to calculate the harmonic mean of the number of replications:

$$
p_h = \frac{n}{\sum^n_{i=1} \frac{1}{p_i}}
$$

where $p_h$ is the harmonic mean, $n$ is the number of genotypes, and $p_i$ is the number of replicates of the *i*th genotype.

Now calculate heritability across all environments



```{r calc.herit.all}

# Extract the full model and calculate heritability
stage_two_herit <- stage_two_model_fits %>% 
  mutate(full_fit = map(fits, "full")) %>%
  do(suppressWarnings(herit_boot(object = .$full_fit[[1]], n_e = .$n_e, n_r = .$n_r, boot.reps = 10,
                       exp = "line_name / (line_name + (line_name:env / n_e) + (Residual / n_r))")))

stage_two_herit
    

# Plot
g_herit <- stage_two_herit %>% 
  mutate(herit_correct = heritability - bias) %>%
  ggplot(aes(x = trait, y = herit_correct, ymin = ci_lower, ymax = ci_upper)) +
  geom_col(aes(fill = "fill"), width = 0.80) +
  geom_errorbar(width = 0.5) +
  scale_fill_discrete(guide = FALSE) +
  ylab("Heritability") +
  xlab("Trait") +
  labs(title = "Broad-Sense Heritability",
       caption = "Estimates have been corrected for bias and error bars reflect\na 95% confidence interval of 500 bootstrap replications.") +
  theme_bw()

save_file <- file.path(fig_dir, "heritability.jpg")
ggsave(filename = save_file, plot = g_herit, height = 5, width = 4)
  
```
  

What is the proportion of each variance component to the total phenotypic variance?

```{r var.prop}

prop_varcomp <- stage_two_model_fits %>% 
  mutate(full_fit = map(fits, "full")) %>% 
  do(var_comp = as.data.frame(VarCorr(.$full_fit[[1]]))) %>% 
  unnest() %>%
  select(trait:grp, vcov) %>% 
  group_by(trait) %>%
  mutate(grp = str_replace_all(grp, c("line_name:env" = "GxE", "line_name" = "Genotype", "env" = "Environment")),
         grp = factor(grp, levels = c("Genotype", "GxE", "Environment", "Residual")),
         prop_vcov = vcov / sum(vcov))

# Plot
g_prop_varcomp <- prop_varcomp %>%
  ggplot(aes(x = trait, y = prop_vcov, fill = grp)) +
  geom_col() +
  ylab("Proportion") +
  xlab("Trait") + 
  scale_fill_discrete(guide = guide_legend(title = "Variance\nComponent")) +
  labs(title = "Components of Phenotypic Variance") +
  theme_bw()


save_file <- file.path(fig_dir, "var_comp_prop.jpg")
ggsave(filename = save_file, plot = g_prop_varcomp, height = 5, width = 5)

```



